The Etiology of Cybersecurity

### 2022-04-26 reports/cimss_2022/main-cut.tex
fix oliviero's feedback

\paragraph{Contributions.} 
The goal of this paper is to lay the foundations of a scientific cybersecurity
theory.  We consider the problem raised by Herley not confined to ``computer
security'' but rather we reason on any abstract system (so that our scientific
hypothesis\footnote{In the remainder of this paper, we will use the word
hypothesis to refer to ``scientific hypothesis'' as a proposed scientific
theory that has not gone through an extensive series of tests. We
use ``logical theory'' to refer to a set of formal logical axioms.} may
be tested in any sound implementation such as a network, a mechanical, cyber, or
cyber-physical system, or even a single computer or a single device such as a
hard-drive).  
Instead of starting from reasoning about what makes a system secure or insecure,
we reason about what causes insecurities. 
We focus on insecurities only caused by
the exploitation of cybersecurity attacks, and we assume that achieving
cybersecurity means preventing all those attacks from being exploitable or
exploited.  Our hypothesis is that \emph{cybersecurity attacks are only caused by the
presence of errors in the design or implementation of a system} (i.e.
cybersecurity weaknesses).
With our approach, a list of weaknesses emerges from the mathematical
formulation of a system in a framework called $\abftheory$-framework\ON{Qui, direi o "framework called ABF" o "framework called the ABF-framework"}, and
predicts 4 main classes of weaknesses. Those classes of weaknesses\ON{rimuoverei ripetizione di "of weaknesses"} are used to
calculate all the insecurity configurations of all the components of a system,
obtaining a precise estimation of all potential cybersecurity-related risks in
any given system. Our hypothesis can be falsified by means of experiments,
testing if all the predicted weaknesses are present in the system under consideration,
or testing if other (not predicted by our hypothesis) weaknesses are present.
In fact, if any cybersecurity weaknesses were to be found in a system and not
predicted by our hypothesis, the hypothesis could be declared incomplete.  If a
cybersecurity weakness would be predicted by our hypothesis but found to be impossible
to realize, our hypothesis could be declared as wrong.

@@

\section{A Cybersecurity Hypothesis in the $\abftheory$-framework}\label{sec:hypothesis}
To address the problem raised by Herley, we define how to
distinguish between a secure and an insecure system in the following steps 
of the engineering process:
\begin{enumerate}
	\item \emph{System Specification}: the \emph{functional and physical requirements} \ON{proporrei di togliere l'and dal corsivo} are defined.
	\item \emph{Architecture Design}: the specification is structured into \emph{functional and physical architectures}\ON{same as the above}.
	\item \emph{Cybersecurity Risk Assessment}: potential \emph{weaknesses} (errors) and
		\emph{cybersecurity requirements} are identified.	
\end{enumerate}
We changed the focus from the attacker as the source of insecurity to the
potential design errors of a system. We now define a framework for the
definition of a system that we use to identify weaknesses as potential design
errors.

@@

As in \autocite{Santaca2016abf}, we define agents\ON{so far, we spoke about systems, from now we focus on agents. Shall we briefly motivate why we think of systems in term of agents? Or, at least, say that we use the two terms intercambiabilmente? Later on in the paper, we do say "agent (i.e. system)", but that comes much after this, and so the text inbetween might be puzzling to some.}\ON{Further note: In section 4.1 we explicitly state this. Perhaps shall we here just say "we consider a CPS as a hierarchy of agents, as we will detail later in Section 4.1"? (or something similar)} as a meronomy (an
hierarchy of Part-Whole relations) over the constituents that we previously defined
(assertions, beliefs, and facts), based on a standard definition of mereology,
i.e. based on the definition of parthood relation between \emph{parts}.  Due to
the necessity of considering different relations between parts (as we will show afterwards)
we extend the mereology to a
mereo-topology \autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
considering the relations in Tab.~\ref{tab:rcc358}.  For the sake of
readability, we use the term \emph{region} both to refer to a mereological part
and to a topological region.  Our aim is to create a meronomy (hierarchy of
part-whole relations) instead of the taxonomies (categorization based on
discrete sets) such as the one provided in \autocite{NIST2020NVD,CVE}
so that we don't need to rely on a scoring system (such as the CVSS) to assign
a quantitative evaluation of the cybersecurity of each entry. Instead, we want
a precise calculation of the number of insecure configurations of a system to
emerge from the mathematical formulation of our cybersecurity hypothesis.
A mereotopology, as defined in \autocite{Rachavelpula2017mereotopology},
is a mathematical structure where the basic relation between regions
is the reflexive and symmetric relation \emph{Connects With},
that we use to order a universe of agents
$\agentuniverse$ (see in Tab.~\ref{tab:rcc358}).  We use the Region Connection
Calculus (RCC), as defined in \autocite{bennettLogics,improvingRCC}, to provide
an axiomatization of the mereo-topological concepts. In its broader definition,
the RCC theory is composed by eight axioms, and is known as RCC8 \autocite{Grutter2008rcc}. Using
RCC5 instead of RCC8 prevents us from considering tangential connections
between spatial regions. However, tangential connections in RCC8 can be
considered as special cases of the more general spatial relations considered in
RCC5.
In Tab.~\ref{tab:rcc}, we summarize the axioms of the RCC (see, e.g., \autocite{Grutter2008rcc}).  We can now define a system
over the mereotopology using the RCC calculus, as follows, where $\rcc(X,Y)$ on
two generic regions $X,Y$ represents one of the possible RCC relations between
$X$ and $Y$. We note that all the RCC relations are symmetric with the exception
of those that have an explicit (related) inverse.

\begin{definition}{\bf System State --}\label{def:opsystem}
	A CPS system (or a sub-system) state is defined as the\ON{the->a} tuple
	$s=\langle\rcc(\factRegion,\beliefRegion),\,\rcc(\factRegion,\assertionRegion),\,\rcc(\beliefRegion,\assertionRegion)\rangle$,
	where $\assertionRegion$,$\beliefRegion$, and $\factRegion$ are regions of
	assertions, beliefs (i.e. the beliefs generated by the behavior)\ON{is "generated" the right word here? Anyhow, the bracketed part is not super clear to me.}, and
	facts respectively, expressed as requirements.
\end{definition}

\ON{Shouldn't we introduce these terms here? A, B, F? I mean even with short sentences: assertions are the exchanged data, facts requirements, beliefs are the behaviour...}

As in \autocite{Santaca2016abf}, it follows that, by defining
a system with \ON{with->as?} a fixed number of regions, there exists
an upper-bound to the number of possible configuration of a system, defined by
the possible relations between the different regions.
For completeness, we report in Appendix~\ref{app:agcalc}\ON{here}
some details on the calculation done by the authors in \autocite{Santaca2016abf}.
The general formula to calculate the number of different types of agents is
$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
between $n$ different regions. In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
($\assertionRegion$, $\beliefRegion$, and $\factRegion$), and all the relations
considered in the RCC are binary.  Hence, we have up to 125
different types of agents but only 54 of the 125 (as showed in
\autocite{improvingRCC}) combinations are topologically correct. 
Generalizing to all the RCCs:

@@

%Hence, even if considering a different number of regions than the three
%$\assertionRegion$, $\beliefRegion$, and $\factRegion$ exponentially affects
%the number of theoretical agents, the application of RCC downscales that number
%of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
%3.6 times more (different) types of agents than RCC3, but using RCC8 would
%allow us to consider 3.5 times more different agents.
In the quantitative evaluation of a single agent, as in Fig.~\ref{fig:quantitative},
we argue that only 1 configuration represents the nominal (expected) behavior 
of the agent while the other configurations are either impossible to 
implement or diverge from the intended nominal behavior.\ON{Here, I would explicitly say that what we mean is that, only when the assertiosn, beliefs and facts are equal, then the system is behaving as excepted. That is, I would explicitly say is that the only nominal state is EQ,EQ,EQ. I mean, it is in the table, but a bit implicit to me.} We note 
that the numbers reported here do not consider the details of the
engineering process and should be considered a limit of an abstract 
representation of the system.

@@

agent (i.e. a system) exist w.r.t. the $\abftheory$-framework (i.e.\ON{e.g.?} $54/125$ in RCC5), a

@@

We first analyze the relations between each pair of
regions (i.e.  $\abf$), and then categorize different
agents as tuples of the three regions.\ON{but we don't actually do this, right? Maybe I missed it. If not, we should remove this sentence. If yes, then we should clarify this. (At least to me :-))}  For the sake of simplicity, soundness is
opposed to non-soundness in the following, however, with the RCC, one should\ON{should->could?}

@@

\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration.} 
By definition, assertions are defined as transfer of information
between two agents.\ON{first time we say this!} An agent has two main categories of assertions,
input and output assertions.  Given an agent $a$ and a collection of asserted
predicates $\Phi$,
the input assertions are those received by $a$ from an agent
$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. With a slight abuse
of notation,\ON{alternative: in the text we drop the $\Phi$ when the content of the assertions is not relevant.} in the text we simply write $\assertionRegion_{s\rightarrow a}$ when we abstract
away from the content of the assertion (i.e. $\Phi$).
We shall consider
two pairs of regions:
\begin{itemize}
	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\beliefRegion$},
		where the relation between input-assertions and beliefs
		describes the soundness of the execution of the functional
		architecture w.r.t. input elicitation.
		If all the inputs (assertions) are correctly
		handled in the functional specification (beliefs) the
		specification is complete.\ON{Do we perhaps mean "sound"? We used sound above}

@@

\begin{definition}{\bf Input or Output Port --}\label{def:port} 
	A port forwards information from the outside of an agent's boundary to
	the inside (input-port) or vice versa (output-port).  There exist two
	types of ports with the following behavior: an input-port 
	transforms assertions from a sender $s$ ($\rassert{s}{a}$) to beliefs
	of an agent $a$ ($\beliefRegion_a$), while an output-port transforms beliefs
	into assertions.
\end{definition}
The \emph{quality of a port} is determined by the $\rcc$ relation between the
assertions received or sent and the belief, i.e.
$\Rcc{\rassert{s}{a}}{\beliefRegion_a}$ for input-port or
$\Rcc{\rassert{r}{*}}{\beliefRegion_a}$, for output-port (where $*$ is a wildcard for an agent)\ON{should this $r$ be an $a$?}.  A port is, in
fact, syntactic sugar to express the relation between assertions and beliefs.
We note that the definition of an input/output port can
be considered ``secure'', meaning that we implicitly formulated the requirement
that a port always forwards the information without modifying it in any way.
This is assumed since we defined a port as if the RCC equality relation holds
between the input/output assertions and the input/output beliefs. In contrast,
assuming any \emph{other} RCC relation between inputs and outputs of a port can be considered as generating a weaknesses.

\paragraph{Port Weaknesses.} We now apply the $\abftheory$-framework to
list all possible cybersecurity weaknesses of a port. From our definition of ports, the following holds.
\begin{theorem}[Port Weaknesses]
There exist only the following \emph{six}
types of weaknesses, generating six types of insecure port in RCC5:
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{Replace port}, where assertions reach
		the port but are replaced with different and un-related information
		before passing the boundary.
	\item \emph{Drop port}, where assertions reach the
		port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent).
	\item \emph{Insertion port}, where new information is transferred along with 
		the information incoming from a channel, and then sent to
		the recipient (agent).
	\item \emph{Injection port}, where part of the incoming information
		is substituted with new information and transferred to the intended recipient.
	\item \emph{Selective port}, where some information passes the port and
    part is either: (W5.1) Dropped or (WP5.2) Replaced.
	%\begin{enumerate}[start=1, label={W5.\arabic*)}]
	%	\item Dropped, or 
	%	\item Replaced.
	%\end{enumerate}
\end{enumerate}
\end{theorem}

\begin{proof}
An input port is, in the $\abftheory$-framework, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different types
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Let us
	define a function weight $|X|$ such that, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	every region can be written as union of elements of that base. 
	We distinguish between regions that are related to information and regions that
	don't by writing the latter as $\emptyset$ (e.g. a region corresponding to
	an assertion is written $A\neq\emptyset$, while a region corresponding to an empty assertion
	as $A=\emptyset$).\ON{Do we use the weight/$|A|$ thing in the proof? I don't see it... If not, should we get rid of it?}
	\begin{enumerate}
		\item If $\eq{\assertionRegion}{\beliefRegion}$ then either
			$\assertionRegion=\beliefRegion=\emptyset$ (no
			communication) or
			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
			communication).
		\item If $\dr{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
			(we call \emph{insert} the former, \emph{full drop}
			the latter case), or
			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
			called \emph{replace} (i.e. drop and insert).
		\item If $\pp{\assertionRegion}{\beliefRegion}$ then
			$\beliefRegion$ contains and extend $\assertionRegion$
			which we call \emph{insertion}.
		\item If $\ppi{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion$ contains and extend $\beliefRegion$
			which we call \emph{drop} (or selective drop to stress the difference with the full drop).
		\item If $\po{\assertionRegion}{\beliefRegion}$ then a part of
			the $\assertionRegion$ is contained in the
			$\beliefRegion$ which is a combination of
			\emph{selective drop and generation} which we call \emph{injection}.
	\end{enumerate}
\end{proof}

@@

\begin{definition}{\bf Mono-directional Channel (bottom-channel) --}\label{def:monochannel}
	A mono-directional channel between two agents ($s \rightarrow r$) is
	defined as\ON{remove defined as} an agent whose behavior is (dogmatically) defined as: to
	forward any assertion received from $s$ over an input-port, to  the
	output-port where $r$ is listening to.
\end{definition}
The \emph{quality of a mono-directional} channel is defined as the $\rcc$
relation between the assertions of the sender and the ones received by the
receiver, i.e. $\Rcc{\assertionRegion_s}{\assertionRegion_r}$.

\paragraph{Channel Weaknesses.}
Given that a mono-directional bottom-channel is assumed to be perfectly
forwarding any assertion (as we assumed for ports) from its input-port to its
output-port, there is no insecure behavior but only the combination of the
weaknesses of the input and output port; therefore there exist $(7^2)-1=48$
theoretical configurations ($7^2$ because there are $6$ insecure types of port,\ON{refer to Theorem 1?}
plus $1$ secure type, on both input and output side; and we exclude the
configuration with $2$ secure types as input and output, hence the $-1$); where only 44
are possible. For the sake of readability, we report 6 examples in the
following but the proof by exhaustion over all the possible cases is
straightforward.

@@

\section{Number of different configurations of a system.}\label{agcalc}
The general formula to calculate the number of different types of agents is
$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
between $n$ different regions. Therefore, $r^{\binom{n}{k}}$
expresses the number of permutation of $r$
relations over ${\binom{n}{k}}$ elements with repetitions, 
with ${\binom{n}{k}}$ being the number of
$k$-ary combinations of $n$ regions.
In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
considered in the RCC are binary.  Hence, using RCC5 (with five different
spatial relations) over three regions, we can theoretically define up to 125
different type of agents. However, only 54 of the 125 (as showed in
\autocite{improvingRCC}) combinations are topologically correct with respect to
the definition of the relations of RCC5. Generalizing to all the RCCs:

>>>>

\paragraph{Contributions.} 
The goal of this paper is to lay the foundations of a scientific cybersecurity
theory.  We consider the problem raised by Herley not confined to ``computer
security'' but rather we reason on any abstract system (so that our scientific
hypothesis\footnote{In the remainder of this paper, we will use the word
hypothesis to refer to ``scientific hypothesis'' as a proposed scientific
theory that has not gone through an extensive series of tests. We
use ``logical theory'' to refer to a set of formal logical axioms.} may
be tested in any sound implementation such as a network, a mechanical, cyber, or
cyber-physical system, or even a single computer or a single device such as a
hard-drive).  
Instead of starting from reasoning about what makes a system secure or insecure,
we reason about what causes insecurities. 
We focus on insecurities only caused by
the exploitation of cybersecurity attacks, and we assume that achieving
cybersecurity means preventing all those attacks from being exploitable or
exploited.  Our hypothesis is that \emph{cybersecurity attacks are only caused by the
presence of errors in the design or implementation of a system} (i.e.
cybersecurity weaknesses).
With our approach, a list of weaknesses emerges from the mathematical
formulation of a system in a framework called $\abftheory$ (as in Assertions,
Beliefs, Facts in \autocite{Santaca2016abf})\footnote{Intuitively, as we will
see later, assertions correspond to an exchange of
information between agents, beliefs to internal
information considered true by the agent, and facts to
requirements.} predicts 4 main classes of weaknesses. Those classes are used to
calculate all the insecurity configurations of all the components of a system,
obtaining a precise estimation of all potential cybersecurity-related risks in
any given system. Our hypothesis can be falsified by means of experiments,
testing if all the predicted weaknesses are present in the system under consideration,
or testing if other (not predicted by our hypothesis) weaknesses are present.
In fact, if any cybersecurity weaknesses were to be found in a system and not
predicted by our hypothesis, the hypothesis could be declared incomplete.  If a
cybersecurity weakness would be predicted by our hypothesis but found to be impossible
to realize, our hypothesis could be declared as wrong.

@@

\section{A Cybersecurity Hypothesis in the $\abftheory$-framework}\label{sec:hypothesis}
To address the problem raised by Herley, we define how to
distinguish between a secure and an insecure system in the following steps 
of the engineering process:
\begin{enumerate}
  \item \emph{System Specification}: the \emph{functional} and \emph{physical requirements} are defined.
  \item \emph{Architecture Design}: the specification is structured into \emph{functional} and \emph{physical architectures}.
	\item \emph{Cybersecurity Risk Assessment}: potential \emph{weaknesses} (errors) and
		\emph{cybersecurity requirements} are identified.	
\end{enumerate}
We changed the focus from the attacker as the source of insecurity to the
potential design errors of a system. We now define a framework for the
definition of a system that we use to identify weaknesses as potential design
errors.

@@

Following \autocite{Santaca2016abf}, we define a system as a hierarchy of agents. Furthermore, we model an agent as a meronomy (an
hierarchy of Part-Whole relations) over the aforementioned constituents 
(assertions, beliefs, and facts), based on a standard definition of mereology
(i.e. based on the definition of parthood relation between \emph{parts}).  Due to
the necessity of considering different relations between parts (as we will show afterwards)
we extend the mereology to a
mereo-topology \autocite{Smith1996mereotopology,Varzi1994mereotopology,Rachavelpula2017mereotopology},
considering the relations in Tab.~\ref{tab:rcc358}.  For the sake of
readability, we use the term \emph{region} both to refer to a mereological part
and to a topological region.  Our aim is to create a meronomy (hierarchy of
part-whole relations) instead of the taxonomies (categorization based on
discrete sets) such as the one provided in \autocite{NIST2020NVD,CVE}
so that we don't need to rely on a scoring system (such as the CVSS) to assign
a quantitative evaluation of the cybersecurity of each entry. Instead, we want
a precise calculation of the number of insecure configurations of a system to
emerge from the mathematical formulation of our cybersecurity hypothesis.
A mereotopology, as defined in \autocite{Rachavelpula2017mereotopology},
is a mathematical structure where the basic relation between regions
is the reflexive and symmetric relation \emph{Connects With},
that we use to order a universe of agents
$\agentuniverse$ (see in Tab.~\ref{tab:rcc358}).  We use the Region Connection
Calculus (RCC), as defined in \autocite{bennettLogics,improvingRCC}, to provide
an axiomatization of the mereo-topological concepts. In its broader definition,
the RCC theory is composed by eight axioms, and is known as RCC8 \autocite{Grutter2008rcc}. Using
RCC5 instead of RCC8 prevents us from considering tangential connections
between spatial regions. However, tangential connections in RCC8 can be
considered as special cases of the more general spatial relations considered in
RCC5.
In Tab.~\ref{tab:rcc}, we summarize the axioms of the RCC (see, e.g., \autocite{Grutter2008rcc}).  We can now define a system
over the mereotopology using the RCC calculus, as follows, where $\rcc(X,Y)$ on
two generic regions $X,Y$ represents one of the possible RCC relations between
$X$ and $Y$. We note that all the RCC relations are symmetric with the exception
of those that have an explicit (related) inverse.

\begin{definition}{\bf System State --}\label{def:opsystem}
	A CPS system (or a sub-system) state is defined as a tuple
	$s=\langle\rcc(\factRegion,\beliefRegion),\,\rcc(\factRegion,\assertionRegion),\,\rcc(\beliefRegion,\assertionRegion)\rangle$,
	where $\assertionRegion$,$\beliefRegion$, and $\factRegion$ are regions of
	assertions, beliefs (i.e. the beliefs generated by the behavior), and
	facts respectively, expressed as requirements.
\end{definition}

As in \autocite{Santaca2016abf}, it follows that, by defining
a system as a fixed number of regions, there exists
an upper-bound to the number of possible configuration of a system, defined by
the possible relations between the different regions.
For completeness, we report in Appendix~\ref{app:agcalc}
some details on the calculation done by the authors in \autocite{Santaca2016abf}.
The general formula to calculate the number of different types of agents is
$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
between $n$ different regions. In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
($\assertionRegion$, $\beliefRegion$, and $\factRegion$), and all the relations
considered in the RCC are binary.  Hence, we have up to 125
different types of agents but only 54 of the 125 (as showed in
\autocite{improvingRCC}) combinations are topologically correct. 
Generalizing to all the RCCs:

@@

%Hence, even if considering a different number of regions than the three
%$\assertionRegion$, $\beliefRegion$, and $\factRegion$ exponentially affects
%the number of theoretical agents, the application of RCC downscales that number
%of a factor that ranges from 1.8 to 2.5. In addition, using RCC5 we consider
%3.6 times more (different) types of agents than RCC3, but using RCC8 would
%allow us to consider 3.5 times more different agents.
In the quantitative evaluation of a single agent, as in Fig.~\ref{fig:quantitative},
we argue that only 1 configuration represents the nominal (expected) behavior 
of the agent while the other configurations are either impossible to 
implement or diverge from the intended nominal behavior. We note 
that the numbers reported here do not consider the details of the
engineering process and should be considered a limit of an abstract 
representation of the system.

@@

agent (i.e. a system) exist w.r.t. the $\abftheory$-framework (e.g., $54/125$ in RCC5), a

@@

We now analyze the relations between each pair of
regions (i.e., $\abf$). For the sake of simplicity, soundness is
opposed to non-soundness in the following, however, with the RCC, one could

@@

\paragraph{\Rcc{$\assertionRegion$}{$\beliefRegion$} -- Collaboration.} 
By definition, assertions are defined as transfer of information
between two agents. An agent has two main categories of assertions,
input and output assertions.  Given an agent $a$ and a collection of asserted
predicates $\Phi$,
the input assertions are those received by $a$ from an agent
$s$ acting as a sender, $\rassert{s}{a}{\Phi}$; similarly, output assertions
are sent from $a$ to a receiver $r$, $\rassert{a}{r}{\Phi}$. With a slight abuse
of notation, in the text we drop the $\Phi$ when the content of the assertions is not relevant.
We shall consider
two pairs of regions:
\begin{itemize}
	\item \Rcc{$\assertionRegion_{s\rightarrow a}$}{$\beliefRegion$},
		where the relation between input-assertions and beliefs
		describes the soundness of the execution of the functional
		architecture w.r.t. input elicitation.
		If all the inputs (assertions) are correctly
		handled in the functional specification (beliefs) the
		specification is complete.

@@

\begin{definition}{\bf Input or Output Port --}\label{def:port} 
	A port forwards information from the outside of an agent's boundary to
	the inside (input-port) or vice versa (output-port).  There exist two
	types of ports with the following behavior: an input-port 
	transforms assertions from a sender $s$ ($\rassert{s}{a}$) to beliefs
	of an agent $a$ ($\beliefRegion_a$), while an output-port transforms beliefs
	into assertions.
\end{definition}
The \emph{quality of a port} is determined by the $\rcc$ relation between the
assertions received or sent and the belief, i.e.
$\Rcc{\rassert{s}{a}}{\beliefRegion_a}$ for input-port or
$\Rcc{\rassert{a}{r}}{\beliefRegion_a}$, for output-port.  A port is, in
fact, syntactic sugar to express the relation between assertions and beliefs.
We note that the definition of an input/output port can
be considered ``secure'', meaning that we implicitly formulated the requirement
that a port always forwards the information without modifying it in any way.
This is assumed since we defined a port as if the RCC equality relation holds
between the input/output assertions and the input/output beliefs. In contrast,
assuming any \emph{other} RCC relation between inputs and outputs of a port can be considered as generating a weaknesses.

\paragraph{Port Weaknesses.} We now apply the $\abftheory$-framework to
list all possible cybersecurity weaknesses of a port. From our definition of ports, the following holds.
\begin{theorem}[Port Weaknesses]\label{thm:pw}
There exist only the following \emph{six}
types of weaknesses, generating six types of insecure port in RCC5:
\begin{enumerate}[start=1, label={W\arabic*)}]
	\item \emph{Replace port}, where assertions reach
		the port but are replaced with different and un-related information
		before passing the boundary.
	\item \emph{Drop port}, where assertions reach the
		port but do not pass the boundary of the agent (i.e. do not
		become belief of the agent).
	\item \emph{Insertion port}, where new information is transferred along with 
		the information incoming from a channel, and then sent to
		the recipient (agent).
	\item \emph{Injection port}, where part of the incoming information
		is substituted with new information and transferred to the intended recipient.
	\item \emph{Selective port}, where some information passes the port and
    part is either: (W5.1) Dropped or (WP5.2) Replaced.
	%\begin{enumerate}[start=1, label={W5.\arabic*)}]
	%	\item Dropped, or 
	%	\item Replaced.
	%\end{enumerate}
\end{enumerate}
\end{theorem}

\begin{proof}
An input port is, in the $\abftheory$-framework, defined secure as long as the relation
	between the two regions of input assertions $\assertionRegion$ and
	output beliefs $\beliefRegion$ are equal, i.e.
	$\eq{\assertionRegion}{\beliefRegion}$. Therefore, any other relation
	should result in a weakness (related to an insecurity flaw) of that
	input port.  Using RCC5, there exist exactly other $4$ different types
	of relations, one of which is the discrete-from (DR) relation, i.e.
	$\dr{\assertionRegion}{\beliefRegion}$. When two regions are related by
	the DR relations, they have no subregion in common. Let us
	define a function weight $|X|$ such that, for any region $X$, it
	represents the smallest possible cardinality of a (mereo)topological base for
	$X$; where a base is a collection of regions in a (mereo)topology such that
	every region can be written as union of elements of that base. 
	We distinguish between regions that are related to information and regions that
  are not (i.e., regions $A$ such that $|A|=0$) by writing the latter as $\emptyset$.

  \begin{enumerate}
		\item If $\eq{\assertionRegion}{\beliefRegion}$ then either
			$\assertionRegion=\beliefRegion=\emptyset$ (no
			communication) or
			$\assertionRegion=\beliefRegion\neq\emptyset$ (forward
			communication).
		\item If $\dr{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion=\emptyset\oplus\beliefRegion=\emptyset$
			(we call \emph{insert} the former, \emph{full drop}
			the latter case), or
			$\assertionRegion\neq\emptyset\wedge\beliefRegion\neq\emptyset\wedge\assertionRegion\neq\beliefRegion$
			called \emph{replace} (i.e. drop and insert).
		\item If $\pp{\assertionRegion}{\beliefRegion}$ then
			$\beliefRegion$ contains and extend $\assertionRegion$
			which we call \emph{insertion}.
		\item If $\ppi{\assertionRegion}{\beliefRegion}$ then
			$\assertionRegion$ contains and extend $\beliefRegion$
			which we call \emph{drop} (or selective drop to stress the difference with the full drop).
		\item If $\po{\assertionRegion}{\beliefRegion}$ then a part of
			the $\assertionRegion$ is contained in the
			$\beliefRegion$ which is a combination of
			\emph{selective drop and generation} which we call \emph{injection}.
	\end{enumerate}
\end{proof}

@@

\begin{definition}{\bf Mono-directional Channel (bottom-channel) --}\label{def:monochannel}
	A mono-directional channel between two agents ($s \rightarrow r$) is
	an agent whose behavior is defined as: to
	forward any assertion received from $s$ over an input-port, to  the
	output-port where $r$ is listening to.
\end{definition}
The \emph{quality of a mono-directional} channel is defined as the $\rcc$
relation between the assertions of the sender and the ones received by the
receiver, i.e. $\Rcc{\assertionRegion_s}{\assertionRegion_r}$.

\paragraph{Channel Weaknesses.}
Given that a mono-directional bottom-channel is assumed to be perfectly
forwarding any assertion (as we assumed for ports) from its input-port to its
output-port, there is no insecure behavior but only the combination of the
weaknesses of the input and output port; therefore there exist $(7^2)-1=48$
theoretical configurations ($7^2$ because there are $6$ insecure types of port -- see Theorem \ref{thm:pw} -- plus $1$ secure type, on both input and output side; and we exclude the
configuration with $2$ secure types as input and output, hence the $-1$); where only 44
are possible. For the sake of readability, we report 6 examples in the
following but the proof by exhaustion over all the possible cases is
straightforward.

@@

\section{Number of different configurations of a system.}\label{app:agcalc}
The general formula to calculate the number of different types of agents is
$r^{\binom{n}{k}}$, where $r$ is the number of relations with arity $k$,
between $n$ different regions. Therefore, $r^{\binom{n}{k}}$
expresses the number of permutation of $r$
relations over ${\binom{n}{k}}$ elements with repetitions, 
with ${\binom{n}{k}}$ being the number of
$k$-ary combinations of $n$ regions.
In our case, $\binom{n}{k}=3$ since we consider $3$ regions 
($\assertionRegion,\beliefRegion,\factRegion$), and all the relations
considered in the RCC are binary.  Hence, using RCC5 (with five different
spatial relations) over three regions, we can theoretically define up to 125
different type of agents. However, only 54 of the 125 (as showed in
\autocite{improvingRCC}) combinations are topologically correct with respect to
the definition of the relations of RCC5. Generalizing to all the RCCs:
